---
title: 'Receptor Occupancy Tumor Response'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This repository contains materials for a hands on example for the Effective Visual Communication Webinar. 

* [Purpose_Worksheet.docx](Purpose_Worksheet.docx) is a partially completed Purpose Worksheet setting up the purpose for the visualization exercise.
* [Receptor_Occupancy_Tumor_Response.csv](Receptor_Occupancy_Tumor_Response.csv) is a dataset that can be used to generate visualizations for the exercise.

# Background

This example is inspired by an early oncology clinical trial. Receptor occupancy, tumor size, and RECIST criteria have been altered and no longer reflect the properties of a real drug or trial.

Situation: 

*  Results are available from an early oncology drug trial showing tumor shrinkage and response in a subset of patients
*  Data available include average receptor occupancy (a function of drug exposure)
*  Primary endpoint was overall response (partial or complete response according to RECIST criteria)
*  Secondary endpoint was percent change from baseline tumor diameter 

Complication: 

*  Trial had no control / placebo arm, so not immediately clear whether the tumor shrinkage and response rates are related to treatment 

Question:

*  Is there a way to show the drug is working?

# Purpose

The Purpose Worksheet is intended to help with Step 1 of Effective Visual Communication (i.e., have a clear purpose). For this exercise, we have partially completed the [Purpose_Worksheet](Purpose_Worksheet.docx) for you, based on the background information above. Review the partially completed worksheet, make any changes you wish, and complete what hasn't been filled in.

# Data

A simulated dataset has been provided for this example: [Receptor_Occupancy_Tumor_Response.csv](Receptor_Occupancy_Tumor_Response.csv).  Step 2 of Effective Visual Communication involves you choosing what data to display and how, in order to support the Purpose. You may or may not use all of the information supplied in the dataset.

The dataset represents a set of individual patients following treatment with the oncology drug. The dataset contains individual average receptor occupancy information (a function of drug exposure), as well as individual tumor size and RECIST criteria. Below are descriptions of the variables in the dataset

*  **ID** - unique subject identifier
*  **RO_AVG** - average receptor occupancy (%)
*  **Tumor_PCFBSL** - percent change from baseline in tumor diameter (%)
*  **RECIST** - RECIST criteria ("PD", "SD", "PR", "CR")
*  **RECIST_LABEL** - RECIST criteria descriptor ("PD - Progressive Disease", "SD - Stable Disease", "PR - Partial Response", "CR - Complete Response")

Below find the first six rows of the dataset. Access the full dataset here: [Receptor_Occupancy_Tumor_Response.csv](Receptor_Occupancy_Tumor_Response.csv)

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyr)
library(dplyr)
data <- read.csv("Receptor_Occupancy_Tumor_Response.csv")
knitr::kable(data %>% head(), rownames = FALSE) # %>% mutate(avgAUC = signif(avgAUC, 5), avgAUEC = signif(avgAUEC, 5), HRR = signif(HRR, 5)) %>% head(), rownames = FALSE)
```









